Lattice Boltzmann study of mass transfer for two-dimensional Bretherton/Taylor bubble train flow

### 2012-12-22 Text/Final/answers.tex
Did answers.tex

\begin{document}
It is our pleasure to address the reviewers' comments which improve readability of the manuscript.  Please consider our answers below. All corrections in the manuscript are marked by red.
\section{Reviewer I}

@@

\textbf{Answer:}  We added superscript "LB" which stands for the lattice Boltzmann to distinguish non-dimensional numbers from physical ones in Section 3.1 where the lattice Boltzmann method is introduced. After Section 3.1 all simulations are done in the lattice Boltzmann system, which is non-dimensional. Hereafter, for the sake of simplicity we omit the index. Unless, it is specifically specified, all the quantities after Section 3.1 are non-dimensional (corresponding sentence was added). Thus, the quantities on pages 22 and 25 are non-dimensional. The quantity on page 26 is a physical quantity as it is absolutely rightly specified by referee (corresponding sentence is added).

@@

\textbf{Answer:} We added the values of the corresponding Schmidt numbers to Table 2. For Section 5.1, the diffusion is increased at the same time as velocity increases. Thus, the Schmidt number stays the same  $Sc\approx790$ for all simulations. The value was chosen corresponding to a dye with the low molecular diffusivity in order to avoid diffusion of the dye in the liquid slug and to study mass transfer related to hydrodynamics patterns. The corresponding sentences explaining the influence of the Schmidt number on the numerical simulations are added.

@@

\textbf{Answer:} We rewrote a conclusion to outline better results and its accuracy.

@@

\textbf{Answer:} Thank you for excellent references. We added additional citations and corrected the reference.

@@

\textbf{Answer:}  Thank you for your comments. The amount of simulations we put into this work is large. Especially for the lattice Boltzmann method, being a relatively new method in comparison with all others, many concepts and guidelines of how to perform simulations are yet to be established. Moreover, authors (Onea, et al. Cat Today 147S 2009) who used the VOF method argue that there usage of one periodic cell combined with periodic boundary conditions is questionable. That's why we felt strongly that there is a space for the method papers combined with the guidelines for the lattice Boltzmann method.

@@

\textbf{Answer:} Thank you for your comment.  In the multiphase literature,  some parameters maps, for example flow patterns map, are indicated through superficial velocities of gas and liquid varying as X and Y axes. This is the reason why we wanted to comply. The connection of superficial velocities and gas holdup happens in the equation for the net velocity is $U=U_{\mathrm{liq}} + U_{\mathrm{gas}}=U_{\mathrm{liq,actual}}*\epsilon+U_{\mathrm{gas,actual}}$. However, the actual gas and liquid velocities are connected through the complicated relationship $U_{\mathrm{gas,actual}}=U_{\mathrm{liq,actual}}(Ca,\epsilon)$ as indicated in our previous work and largely depends on the geometry. We added corresponding sentences to the text.

@@

\textbf{Answer:} Thank you for noticing it. The typo is corrected. We used macros and the error was in the definition and propagated to Eq.12 and 13.

@@

\textbf{Answer:} Thank you for corrections. We corrected the sign before normal component and added the better explanation in Appendix. In fact, the free slip is indeed simulated  - one can see the streamlines in the reference frame moving with the bubble. Those streamlines patterns coincide within $3\%$ with the streamlines already validated in our previous works \cite{kuzmin-binary2d,kuzmin-binary3d}.

@@

\textbf{Answer:} We rewrote the caption text to the figure to explain better. All the lines are close to each other and that's the thought we want to deliver. If some lines are laying apart then the simulations produce different results, thus scaling procedure is not consistent.

@@

\textbf{Answer:} Thank you for improving readability of the article. We wanted to say that parameters indicated in Table 2  give $Pe=14041$. The velocity due to scaling is around $U\approx 0.1$ and that means to achieve such parameter we need to have extremely low (for lattice Boltzmann simulations) diffusion coefficient $D=U L / Pe$. The diffusion parameter $D$ is connected with the lattice Boltzmann relaxation parameter $\omega$. For low values of diffusion the relaxation parameter $\omega\approx 1.99$ which is close to theoretical stability limit $\omega_{crit}=2$. In simulations we see that simulations are unstable with the velocity pattern for $Ca=1.040$ and $Pe=14041$. This is the reason that we decrease Peclet number (or $Sc=794.84/5=158.96$)  by increasing diffusion parameter $D$ by $5$ times. The high $Pe$ and $Sc$ numbers simulations are difficult to perform not only with the LBM, but with the Volume of Fluid Method as well \cite{onea-mass}. The corresponding sentences with explanations are added.

@@

\textbf{Question:} Table 1 - I did not find what $epsilon_gas$ means. Is it the gas holdup?

>>>>

\begin{document}
It is our pleasure to address the reviewers' comments which improved the manuscript.  Please consider our answers below. All corrections in the manuscript are marked by red.
\section{Reviewer I}

@@

\textbf{Answer:}  We added superscript "LB" which stands for the lattice Boltzmann to distinguish non-dimensional numbers from physical ones in Section 3.1 where the lattice Boltzmann method is introduced. After Section 3.1 all simulations are done in the lattice Boltzmann system, which is non-dimensional. Hereafter, for the sake of simplicity we omit the index. Unless, it is specifically specified, all the quantities after Section 3.1 are non-dimensional (corresponding sentence was added). Thus, the quantities on pages 22 and 25 are non-dimensional. The quantity on page 26 is a physical quantity as it is absolutely rightly specified by the referee (corresponding sentence is added).

@@

\textbf{Answer:} We added the values of the corresponding Schmidt numbers to Table 2. For Section 5.1, the diffusion is increased with the same factor as  the velocity increases. Thus, the Schmidt number stays the same  $Sc\approx790$ for all simulations. The value was chosen corresponding to a dye with the low molecular diffusivity in order to avoid diffusion of the dye in the liquid slug and to study mass transfer related to hydrodynamics patterns. The corresponding sentences explaining the influence of the Schmidt number on the numerical simulations are added.

@@

\textbf{Answer:} We rewrote a conclusion to outline better results and their accuracy.

@@

\textbf{Answer:} Thank you for excellent references. We added the additional citations and corrected the reference.

@@

\textbf{Answer:}  Thank you for your comments. The amount of simulations we put into this work is large. Especially for the lattice Boltzmann method, being a relatively new method in comparison with all others, many concepts and guidelines of how to perform simulations are yet to be established. Moreover, authors (Onea, et al. Cat Today 147S 2009) who used the VOF method argue that their usage of one periodic cell combined with periodic boundary conditions is questionable. That's why we felt strongly that there is space for a  methods paper combined with the guidelines for the lattice Boltzmann method.

@@

\textbf{Answer:} Thank you for your comment.  In the multiphase literature,  some parameters maps, for example flow patterns map, are indicated through superficial velocities of gas and liquid varying as X and Y axes. This is the reason why we wanted to comply. The connection of superficial velocities and gas holdup happens in the equation for the net velocity is $U=U_{\mathrm{liq}} + U_{\mathrm{gas}}=U_{\mathrm{liq,actual}}*\epsilon+U_{\mathrm{gas,actual}} (1-\epsilon)$. Thus, the gas holdup is already included in the equation. Note that the actual gas and liquid velocities are connected through the complicated relationship $U_{\mathrm{gas,actual}}=U_{\mathrm{liq,actual}}(Ca,\epsilon)$ as indicated in our previous work and largely depends on the geometry. We added corresponding sentences to the text.

@@

\textbf{Answer:} Thank you for noticing it. The typo is corrected.

@@

\textbf{Answer:} Thank you for corrections. We corrected the sign before the normal component and added a better explanation in  the Appendix. In fact, the free slip is indeed simulated  - one can see the streamlines in the reference frame moving with the bubble. Those streamlines patterns coincide within $3\%$ with the streamlines already validated in our previous works \cite{kuzmin-binary2d,kuzmin-binary3d}.

@@

\textbf{Answer:} We rewrote the caption text to the figure to explain better. All the lines are close to each other and that's the message we want to deliver. If some lines are laying apart then the simulations produce different results, and the scaling procedure is not consistent.

@@

\textbf{Answer:} Thank you for improving the readability of the article. We wanted to say that parameters indicated in Table 2  give $Pe=14041$. The velocity due to scaling is around $U\approx 0.1$ and that means to achieve such parameter we need to have extremely low (for lattice Boltzmann simulations) diffusion coefficient $D=U L / Pe$. The diffusion parameter $D$ is connected with the lattice Boltzmann relaxation parameter $\omega$. For low values of diffusion the relaxation parameter $\omega\approx 1.99$ which is close to theoretical stability limit $\omega_{crit}=2$. In simulations we see that simulations are unstable with the velocity pattern for $Ca=1.040$ and $Pe=14041$. This is the reason that we decrease the Peclet number (or $Sc=794.84/5=158.96$)  by increasing the diffusion parameter $D$ by $5$ times. The high $Pe$ and $Sc$ numbers simulations are difficult to perform not only with the LBM, but with the Volume of Fluid Method as well \cite{onea-mass}. The corresponding sentences with explanations are added.

@@

\textbf{Question:} Table 1 - I did not find what $\epsilon_{gas}$ means. Is it the gas holdup?

### 2012-12-22 Text/Final/introduction.tex
Did answers.tex

\begin{document}
\begin{frontmatter}
\title{Lattice Boltzmann study of mass transfer for two-dimensional Bretherton/Taylor bubble train flow}
\author[uofa]{A.~Kuzmin\corref{cor1}}
\cortext[cor1]{Corresponding author, Telephone: +1(438)824-3695}
\ead{kuzmin@ualberta.ca}
\author[us]{M.~Januszewski}
\ead{michalj@gmail.com}
\author[schlum]{D.~Eskin}
\ead{deskin@slb.com}
\author[schlum]{F.~Mostowfi}
\ead{fmostowfi@slb.com}
\author[uofa]{J.J.~Derksen}
\ead{jos@ualberta.ca}
\address[uofa]{Chemical and Materials Engineering, University of Alberta\\ 7th Floor, ECERF, 9107
116 St, Edmonton, Alberta, T6G
2V4 Canada}
\address[us]{Insitute of Physics, University of Silesia, 40-007 Katowice, Poland}
\address[schlum]{Schlumberger DBR Technology Center\\ 9450 17 Ave NW, Edmonton, Alberta, T6N 1M9
Canada}
%\address[google]{Google Switzerland GmbH, Brandschenkestrasse 110, 8002 Zurich, Switzerland}
\begin{abstract}
This work presents a procedure for the determination of the volumetric mass transfer
coefficient in the context of lattice Boltzmann simulations for the Bretherton/Taylor bubble train
flow for capillary numbers $0.1 < Ca < 1.0$. We address the case where the hydrodynamic pattern changes from having a
vortex in a slug ($Ca<0.7$) to not having it ($Ca>0.7$) \cite{giavedoni-numerical}.  In the latter case
the bubble shape is asymmetric and cannot be approximated through flat surfaces and circular circumferences as
is often done in the literature \cite{vanbaten-circular,kreutzer-overview}.  When the vortex is present
in the slug, the scalar concentration is well mixed and it is common to use periodic boundary conditions and the inlet/outlet-averaged concentration
% Do we really want equation references in the abstract?
as the characteristic concentration. %, Eq. \ref{eq:main:definition}.
The latter is not valid for flows where the tracer is not well mixed, i.e. $Ca>0.7$.  We therefore
examine various boundary conditions (periodic, open, open with more than 1 unit cell) and definitions of the characteristic concentration to estimate mass transfer coefficients  for the range of capillary numbers $0.1<Ca<1.0$.
We show that the time-dependent volume averaged concentration taken as the characteristic concentration produces the most robust results and
that all strategies presented in the literature are extreme limits of one unified equation. %,
% Same question applies.
%Eq. \ref{main:main:main}. 
Finally, we show good agreement of simulation results for different Peclet numbers
with analytical predictions of \citet{vanbaten-circular}. %The work is of use to people performing mass transfer numerical simulations for bubble train flow in microchannels within the lattice Boltzmann context.
\end{abstract}
\begin{keyword}
Mass Transfer \sep Taylor/Bretherton bubble train flow \sep Multiphase flow \sep Lattice Boltzmann
method
\sep Binary liquid model \sep Flow in microchannels with parallel plates 
%% keywords here, in the form: keyword \sep keyword

@@

\section{Introduction}
\label{intro}
Monolith reactors have recently been getting more attention as a promising alternative to slurry
reactors and trickle bed reactors \cite{kreutzer-overview,bercic-mass}.  These reactors usually operate in
 the Bretherton-Taylor regime \cite{bretherton,taylor} which is a flow
 of equally sized, long air bubbles through a liquid medium, see
Fig. \ref{fig:benchmark:hydro}. This flow regime
is characterized by the dominance of surface tension over inertia and viscous effects, and by
comparatively small gas flow velocities \cite{yue-mass}. Due to the dominance of surface tension, the flow exhibits advantageous
 properties which cannot be achieved
in its macroscopic counterparts: liquid thin films \cite{bretherton} between bubbles and walls strongly enhance mass
transfer from gas and walls to liquid; the plug flow regime occurring in monolith reactors allows to perform chemical reactions in slugs only \cite{kreutzer-overview}.
 Moreover, the low slip velocity between gas and liquid is utilized in experiments to measure liquid velocity \cite{taylor}:  bubbles
 travelling with approximately the same velocity as liquid can be captured with a camera. These properties explain why nowadays one can find a large number of applications
of the Bretherton-Taylor bubble train flow: continuous flow analyzers to measure liquid velocity, chemical reactors for hydrogenation of nitroaromatics, 2-ethyl-hexenal, Fischer-Tropsch synthesis, etc. The extensive reviews of \citet{kreutzer-overview,
gupta-review,yue-mass} cover a significant number of applications.

>>>>

\begin{document}
\begin{frontmatter}
\title{Lattice Boltzmann study of mass transfer for two-dimensional Bretherton/Taylor bubble train flow}
\author[uofa]{A.~Kuzmin\corref{cor1}}
\cortext[cor1]{Corresponding author, Telephone: +1(438)824-3695}
\ead{kuzmin@ualberta.ca}
\author[us]{M.~Januszewski}
\ead{michalj@gmail.com}
\author[schlum]{D.~Eskin}
\ead{deskin@slb.com}
\author[schlum]{F.~Mostowfi}
\ead{fmostowfi@slb.com}
\author[uofa]{J.J.~Derksen}
\ead{jos@ualberta.ca}
\address[uofa]{Chemical and Materials Engineering, University of Alberta\\ 7th Floor, ECERF, 9107
116 St, Edmonton, Alberta, T6G
2V4 Canada}
\address[us]{Insitute of Physics, University of Silesia, 40-007 Katowice, Poland}
\address[schlum]{Schlumberger DBR Technology Center\\ 9450 17 Ave NW, Edmonton, Alberta, T6N 1M9
Canada}
%\address[google]{Google Switzerland GmbH, Brandschenkestrasse 110, 8002 Zurich, Switzerland}
\begin{abstract}
This work presents a procedure for the determination of the volumetric mass transfer
coefficient in the context of lattice Boltzmann simulations for the Bretherton/Taylor bubble train
flow for capillary numbers $0.1 < Ca < 1.0$. We address the case where the hydrodynamic pattern changes from having a
vortex in the slug ($Ca<0.7$) to not having it ($Ca>0.7$) \cite{giavedoni-numerical}.  In the latter case
the bubble shape is asymmetric and cannot be approximated through flat surfaces and circular circumferences as
is often done in the literature \cite{vanbaten-circular,kreutzer-overview}.  When the vortex is present
in the slug, the scalar concentration is well mixed and it is common to use periodic boundary conditions and the inlet/outlet-averaged concentration
% Do we really want equation references in the abstract?
as the characteristic concentration. %, Eq. \ref{eq:main:definition}.
The latter is not valid for flows where the tracer is not well mixed, i.e. $Ca>0.7$.  We therefore
examine various boundary conditions (periodic, open, open with more than 1 unit cell) and definitions of the characteristic concentration to estimate mass transfer coefficients  for the range of capillary numbers $0.1<Ca<1.0$.
We show that the time-dependent volume averaged concentration taken as the characteristic concentration produces the most robust results and
that all strategies presented in the literature are extreme limits of one unified equation. %,
% Same question applies.
%Eq. \ref{main:main:main}. 
Finally, we show good agreement of simulation results for different Peclet numbers
with analytical predictions of \citet{vanbaten-circular}. %The work is of use to people performing mass transfer numerical simulations for bubble train flow in microchannels within the lattice Boltzmann context.
\end{abstract}
\begin{keyword}
Mass Transfer \sep Taylor/Bretherton bubble train flow \sep Multiphase flow \sep Lattice Boltzmann
method
\sep Binary liquid model \sep Flow in microchannels with parallel plates 
%% keywords here, in the form: keyword \sep keyword

@@

\section{Introduction}
\label{intro}
Monolith reactors have recently been getting more attention as a promising alternative to slurry
reactors and trickle bed reactors \cite{kreutzer-overview,bercic-mass}.  These reactors usually operate in
 the Bretherton-Taylor regime \cite{bretherton,taylor} which is a flow
 of equally sized, long air bubbles through a liquid medium, see
Fig. \ref{fig:benchmark:hydro}. This flow regime
is characterized by the dominance of surface tension over inertia and viscous effects, and by
comparatively small gas flow velocities \cite{yue-mass}. Due to the dominance of surface tension, the flow exhibits advantageous
 properties which cannot be achieved
in its macroscopic counterparts: thin liquid films \cite{bretherton} between bubbles and walls strongly enhance mass
transfer from gas and walls to liquid; the plug flow regime occurring in monolith reactors allows to perform chemical reactions in slugs only \cite{kreutzer-overview}.
 Moreover, the low slip velocity between gas and liquid is utilized in experiments to measure liquid velocity \cite{taylor}:  bubbles
 travelling with approximately the same velocity as liquid can be captured with a camera. These properties explain why nowadays one can find a large number of applications
of the Bretherton-Taylor bubble train flow: continuous flow analyzers to measure liquid velocity, chemical reactors for hydrogenation of nitroaromatics, 2-ethyl-hexenal, Fischer-Tropsch synthesis, etc. The extensive reviews of \citet{kreutzer-overview,
gupta-review,yue-mass} cover a significant number of applications.
